Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries

### 2022-03-04 Manuscript/manuscript_text.tex
Updated manuscript

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance. Scenario 1: no effect of boundary. Scenario 2: negative edge effect on yield. Scenario 3: positive ecosystem service effect on yield. Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
\end{figure}

@@

Precision yield data were collected directly from farmers across Alberta.
Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
We converted raw data to a standard csv format using Ag Leader SMS®.
85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
Individual fields ranged from 5.9 to 369 ha in area (mean: 72.2 ha), and contained between 1 and 6 years of data (mean: 3.1), for a total of 14.4 million data points.

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/Field Locations} \caption{Locations of fields that data were collected from}\label{fig:fieldLocataions}
\end{figure}

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):

@@

In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).

@@

The filtered data from each field contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
Figure \ref{fig:spatialSmooths}a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths}b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field. Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field. Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.}\label{fig:spatialSmooths}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).}\label{fig:distSmooths}
\end{figure}

@@

Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Boxplot of field boundary compositions across sampled fields. The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.}\label{fig:boundaryTypes}
\end{figure}

@@

\begin{table}

\caption{\label{tab:smoothSummary}Mean $\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r}
\hline
\multicolumn{1}{c|}{} & \multicolumn{3}{c|}{Mean Yield} & \multicolumn{3}{c}{Yield Variability} \\
\cline{2-4} \cline{5-7}
Smoother Type & $\chi^2$-value & p-value & Prop. <0.05 & $\chi^2$-value & p-value & Prop. <0.05\\
\hline
Bare & 154.7 & 0.036 & 0.877 & 121.4 & 0.054 & 0.860\\
\hline
Grassland & 303.1 & 0.079 & 0.794 & 93.2 & 0.061 & 0.794\\
\hline
Other Crop & 174.9 & 0.063 & 0.860 & 104.5 & 0.044 & 0.840\\
\hline
Shelterbelt & 328.1 & 0.014 & 0.936 & 137.6 & 0.058 & 0.817\\
\hline
Standard & 281.9 & 0.021 & 0.938 & 159.6 & 0.057 & 0.832\\
\hline
Wetland & 148.9 & 0.045 & 0.833 & 106.7 & 0.013 & 0.857\\
\hline
Spatial Smoother & 17357.5 & 0.000 & 1.000 & 4739.1 & 0.000 & 1.000\\
\hline
\end{tabular}
\end{table}

Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of spatial variation. Upper panel represents mean yield, while the lower panel represents yield variation. N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:canolaPlot}
\end{figure}

The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
Average yield (Figure \ref{fig:wheatPlot}, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure \ref{fig:wheatPlot}, bottom row).

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of spatial variation. Upper panel represents mean yield, while the lower panel represents yield variation. N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:wheatPlot}
\end{figure}

Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
Conversely, yield variability (Figure \ref{fig:peaPlot}, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.

\begin{figure}
\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of spatial variation. Upper panel represents mean yield, while the lower panel represents yield variation. N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:peaPlot}
\end{figure}

@@

Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.

@@

Our first scenario (Scenario 1, Figure \ref{fig:hypotheses}) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
Lastly, pest pressure or pollination requirements may be low enough that ecosystem service provision does not significantly change yield.
Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
Crop pests in most western Canadian fields are suppressed by insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty, meaning spillover of beneficial predators and parasitoids may only matter during outbreak years.

Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield.
Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
Between-year effects may be important, especially during hot dry years (Kort 1988), but this was not considered in our analysis.
Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
Finally, about 50\% of growers in western Canada have collected precision agriculture data, and about 85\% have yield monitoring capability (Steel 2017), which means that our work has only scratched the surface of what could be (collectively) one of the richest agroecological datasets ever created.
We urge agronomists, growers, and ecologists to adopt more wide-spread use of precision yield data, as it represents a vast source of incredibly detailed information which can be used to answer a variety of important questions.

@@

\singlespacing
\nolinenumbers

\newpage

\hypertarget{appendix-a-r-code-for-data-filtering}{%
\section*{Appendix A: R Code for Data Filtering}\label{appendix-a-r-code-for-data-filtering}}
\addcontentsline{toc}{section}{Appendix A: R Code for Data Filtering}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Helper functions {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}

\CommentTok{\#"Inlier" spatial filtering procedure from Vega et al 2019, rewritten to work}
\CommentTok{\#with sf + dplyr. Returns boolean}
\NormalTok{vegaFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data,ycol,}\AttributeTok{pvalCutoff=}\FloatTok{0.05}\NormalTok{,}\AttributeTok{nDist=}\DecValTok{40}\NormalTok{)\{}
  \FunctionTok{require}\NormalTok{(sf)}
  \FunctionTok{require}\NormalTok{(sp)}
  \FunctionTok{require}\NormalTok{(spdep)}
  \FunctionTok{require}\NormalTok{(tidyverse)}

  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{any}\NormalTok{(}\FunctionTok{class}\NormalTok{(data) }\SpecialCharTok{\%in\%} \StringTok{\textquotesingle{}sf\textquotesingle{}}\NormalTok{)) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Dataframe must be sf object\textquotesingle{}}\NormalTok{)}

\NormalTok{  coords }\OtherTok{\textless{}{-}} \FunctionTok{st\_coordinates}\NormalTok{(data) }\CommentTok{\#Get coordinates in matrix form}

  \CommentTok{\#Get neighbourhood weights from 0 to ndist meters}
\NormalTok{  nWeights }\OtherTok{\textless{}{-}} \FunctionTok{dnearneigh}\NormalTok{(coords,}\DecValTok{0}\NormalTok{,nDist) }

  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{any}\NormalTok{(}\FunctionTok{sapply}\NormalTok{(nWeights,length)}\SpecialCharTok{==}\DecValTok{1}\NormalTok{))\{}
    \FunctionTok{warning}\NormalTok{(}\StringTok{\textquotesingle{}Some points had no neighbours and were removed\textquotesingle{}}\NormalTok{)}
\NormalTok{  \} }

  \CommentTok{\#Get neighbourhood indices for each point (which other points are in this}
  \CommentTok{\#point\textquotesingle{}s neighbourhood?)}
\NormalTok{  nIndices }\OtherTok{\textless{}{-}} \FunctionTok{nb2listw}\NormalTok{(nWeights, }\AttributeTok{style =} \StringTok{"W"}\NormalTok{,}\AttributeTok{zero.policy =} \ConstantTok{TRUE}\NormalTok{) }

\NormalTok{  yield }\OtherTok{\textless{}{-}} \FunctionTok{pull}\NormalTok{(data,\{\{ycol\}\}) }\CommentTok{\#Get yield data column}

  \CommentTok{\#Local Moran\textquotesingle{}s I for each neighbourhood}
\NormalTok{  LM }\OtherTok{\textless{}{-}} \FunctionTok{localmoran}\NormalTok{(yield,nIndices,}\AttributeTok{p.adjust.method=}\StringTok{"bonferroni"}\NormalTok{,}
                   \AttributeTok{alternative =}\StringTok{"less"}\NormalTok{)}

\NormalTok{  results }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(LM) }\SpecialCharTok{\%\textgreater{}\%} 
    \FunctionTok{rename}\NormalTok{(}\StringTok{\textquotesingle{}pval\textquotesingle{}}\OtherTok{=}\FunctionTok{contains}\NormalTok{(}\StringTok{\textquotesingle{}Pr.z.\textquotesingle{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
    \CommentTok{\#Filter negative Ii and pvals \textless{} 0.05}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{keepThese=}\NormalTok{ Ii }\SpecialCharTok{\textgreater{}} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ pval }\SpecialCharTok{\textgreater{}}\NormalTok{ pvalCutoff) }\SpecialCharTok{\%\textgreater{}\%} 
    \CommentTok{\#NAs (points had no neighbours)}
    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{keepThese=}\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(keepThese),}\ConstantTok{FALSE}\NormalTok{,keepThese)) }

\NormalTok{  ret }\OtherTok{\textless{}{-}} \FunctionTok{pull}\NormalTok{(results)}
  \FunctionTok{return}\NormalTok{(ret)}
\NormalTok{\}}

\CommentTok{\#Function to filter anything above a certain z{-}score. Returns boolean}
\NormalTok{ZscoreFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,}\AttributeTok{z=}\DecValTok{3}\NormalTok{)\{}
\NormalTok{  xmean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(x,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}
\NormalTok{  xsd }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(x,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}
\NormalTok{  zscore }\OtherTok{\textless{}{-}} \FunctionTok{abs}\NormalTok{(((x}\SpecialCharTok{{-}}\NormalTok{xmean)}\SpecialCharTok{/}\NormalTok{xsd))}
\NormalTok{  zscore}\SpecialCharTok{\textless{}}\NormalTok{z}
\NormalTok{\} }

\CommentTok{\#Function to filter anything above certain quantiles. Returns boolean}
\NormalTok{QuantileFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,}\AttributeTok{quant=}\FloatTok{0.99}\NormalTok{)\{ }
\NormalTok{  l }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{((}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{quant)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{,}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{quant)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{) }\CommentTok{\#Symmetric quantiles}
\NormalTok{  x}\SpecialCharTok{\textgreater{}}\FunctionTok{quantile}\NormalTok{(x,l[}\DecValTok{1}\NormalTok{]) }\SpecialCharTok{\&}\NormalTok{ x}\SpecialCharTok{\textless{}}\FunctionTok{quantile}\NormalTok{(x,l[}\DecValTok{2}\NormalTok{]) }
\NormalTok{\}}

\CommentTok{\#Function to filter large changes in bearing. Returns boolean unless}
\CommentTok{\#returnDiffs==TRUE}
\NormalTok{bearingFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(bearing,}\AttributeTok{q=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{z=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{returnDiffs=}\ConstantTok{FALSE}\NormalTok{)\{}
  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{xor}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(q),}\FunctionTok{is.null}\NormalTok{(z))}\SpecialCharTok{\&!}\NormalTok{returnDiffs) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Input quantiles or Z{-}score\textquotesingle{}}\NormalTok{)}
  \CommentTok{\#Difference in compass bearings (in degrees)}
\NormalTok{  bearingDiff }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x1,x2)\{}
\NormalTok{    x }\OtherTok{\textless{}{-}}\NormalTok{ x1}\SpecialCharTok{{-}}\NormalTok{x2}
\NormalTok{    x }\OtherTok{\textless{}{-}} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x)}\SpecialCharTok{\textgreater{}}\DecValTok{180}\NormalTok{,x}\SpecialCharTok{{-}}\NormalTok{(}\DecValTok{360}\SpecialCharTok{*}\FunctionTok{sign}\NormalTok{(x)),x) }\CommentTok{\#Angle differences can\textquotesingle{}t be \textgreater{}180}
    \FunctionTok{return}\NormalTok{(x)}
\NormalTok{  \}}

>>>>

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure 1, Scenario 2).
Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).

@@

Precision yield data were collected directly from farmers across Alberta (Figure S1).
Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
We converted raw data to a standard csv format using Ag Leader SMS®.
85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
Individual fields ranged from 5.9 to 369 ha in area (mean: 72.2 ha), and contained between 1 and 6 years of data (mean: 3.1), for a total of 14.4 million data points.

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):

@@

In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure 1).
All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.38, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.5, Wickham 2016; and 0.4.0, Kassambara 2020).

@@

The filtered data from each field contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
Figure 2a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure 3).
Finally, the random spatial smoothers in Figure 2b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.

@@

Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure S2).

@@

Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure 4 top row).
Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
Yield variability (Figure 4 bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.

The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
Average yield (Figure 5, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure 5, bottom row).

Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
Average pea yields (Figure 6, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
Conversely, yield variability (Figure 6, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.

@@

Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure 1).
There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.

@@

Our first scenario (Scenario 1, Figure 1) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
Lastly, pest pressure or pollination requirements may be low enough that ecosystem service provision does not significantly change yield.
Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
Crop pests in most western Canadian fields are suppressed by insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty, meaning spillover of beneficial predators and parasitoids may only matter during outbreak years.

Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield.
Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure S2), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
Between-year effects may be important, especially during hot dry years (Kort 1988), but this was not considered in our analysis.
Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
Finally, about 50\% of growers in western Canada have collected precision agriculture data, and about 85\% have yield monitoring capability (Steel 2017), which means that our work has only scratched the surface of what could be (collectively) one of the richest agroecological datasets ever created.
We urge agronomists, growers, and ecologists to adopt more wide-spread use of precision yield data, as it represents a vast source of incredibly detailed information which can be used to answer a variety of important questions.

@@
